\vspace{1.0cm} \LARGE \bf Revealing the competitive dynamics of influenza viruses through evolutionary cartography

### 2012-12-19 flux.tex
Finish conclusions section.

In this context, it is especially interesting that adaptive fixations in the HA1 protein determined through analysis of nonsynonymous and synonymous evolution is substantially greater in A/H3N2 than in A/H1N1 \cite{Wolf06, Bhatt11}.
Thus, we conclude that additional adaptive fixations at the amino acid level translate to greater rates of adaptive antigenic change.

@@

In this study we provide a foundation for evolutionary antigenic cartography in which to simultaneously assess antigenic phenotype and patterns of antigenic evolution.
We use this approach to characterize competitive dynamics across influenza clades A/H3N2, A/H1N1, B/Vic and B/Yam and show that antigenic evolution drives strain replacement in each influenza clade.
However, A/H3N2 shows substantially stronger signals of adaptive antigenic evolution, evolving faster in antigenic phenotype and showing more efficient conversion of transient antigenic polymorphism into fixed differences.
Consequently, we observe substantially greater levels of incidence in A/H3N2 than in other influenza clades.
We suggest that antigenic evolution strongly influences competitive dynamics both within and between influenza clades.

>>>>

In this context, it is especially interesting that adaptive fixation in the HA1 protein determined through analysis of nonsynonymous and synonymous evolution is substantially greater in A/H3N2 than in A/H1N1 \cite{Wolf06, Bhatt11}.
Thus, we conclude that additional adaptive fixations at the amino acid level translate to greater rates of adaptive antigenic change.

@@

In this study we provide a foundation for evolutionary antigenic cartography, which seeks to simultaneously assess antigenic phenotype and antigenic evolution.
We use this approach to characterize competitive dynamics across influenza clades A/H3N2, A/H1N1, B/Vic and B/Yam and show that antigenic evolution drives strain replacement in each influenza clade.
However, A/H3N2 shows substantially stronger signals of adaptive antigenic evolution, evolving faster in antigenic phenotype and showing more efficient conversion of transient antigenic polymorphism into fixed differences.
Correspondingly, we observe substantially greater levels of incidence in A/H3N2 than in other influenza clades.
We suggest that antigenic evolution strongly influences competitive dynamics both within and between influenza clades.

The statistical framework presented here represents a baseline to which further advancements in modeling antigenic phenotype and evolution may be made.
For example, our likelihood-based model faciliates the inclusion of possible covariates affecting immunological titer, which could include experimental factors such as red blood cell type is used in the HI assay \cite{Lin12} and whether oseltamivir is included in the HI reaction \cite{Lin10}.
Additionally, this framework should be ideally suited to uncovering genetic determinates of antigenic change, as both the sequence state and antigenic location of internal nodes in the phylogeny may be estimated.
In this fashion, it should be possible to correlate sequence substitutions directly to antigenic diffusion.

Further work characterizing antigenic evolution in the human influenza virus may eventually prove to be crucial in improving vaccine strain selection.
